---
title: "Dissert-final-exp"
author: "Gabriel Robaina"
date: "2024-03-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("main-charts.R")
source("dissert-anova.R")
source("dissert-conf-int.R")
source("dissert-cv.R")
source("dissert-mean-median-diff.R")

options(warn=-1)
```

# Dissert experiments

## Write operations on large files

```{r write-ops-large-files, echo=FALSE}
ecdfMaxFileAndMinMaxIoWrite()
maxFileConfInt(maxFileMinMaxIoWriteQuery())
largeFilesWriteCvBar()
histogramMaxFileMinMaxIoWrite()
bigFileWriteAnova()
```

## Write operations on small files

```{r write-ops-small-files, echo=FALSE}
ecdfMinFileMinIoMinMaxResourceWrite()
minFileConfInt(minFileMinIoMinMaxResourceWriteQuery())
smallFilesWriteCvBar()
histogramMinFileMinMaxResourceTierWrite()
smallFileWriteAnova()
```

## Read operations on large files

```{r read-ops-large-files, echo=FALSE}
ecdfMaxFileAndMinMaxIoRead()
maxFileConfInt(maxFileMinMaxIoReadQuery())
largeFilesReadCvBar()
histogramMaxFileMinMaxIoRead()
bigFileReadAnova()
```

## Read operations on small files

```{r read-ops-small-files, echo=FALSE}
ecdfMinFileMinIoMinMaxResourceRead()
minFileConfInt(minFileMinIoMinMaxResourceReadQuery())
smallFilesReadCvBar()
histogramMinFileMinMaxResourceTierRead()
smallFileReadAnova()
```

## Other plots

The following are [density plots](https://r-graph-gallery.com/density-plot) for experiments ran between 02/25/2024 and 03/20/2024.

On the first density plot, it is limited between 0 and 1. Its hiding a mode for Lambda_DD on Tier 1 that happens after 25 and 50 ms.

### Density plots
```{r densities, echo=FALSE}
densityMinFileAndMinIoWrite()
densityMaxFileAndMinIoWrite()
densityMaxFileAndMaxIoWrite()

densityMinFileAndMinIoRead()
densityMaxFileAndMinIoRead()
densityMaxFileAndMaxIoRead()
```

### Box plots

```{r boxplots, echo=FALSE}
boxplotMinFileAndMinIoWrite()
boxplotMaxFileAndMinIoWrite()
boxplotMaxFileAndMaxIoWrite()

boxplotMinFileAndMinIoRead()
boxplotMaxFileAndMinIoRead()
boxplotMaxFileAndMaxIoRead()
```